---
title: "Plant Community Wrangling"
output: html_document
date: "2023-06-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(ggplot2)


```

```{r}
survey_1A_2023_04_12 = read_csv(here("data/2023/plant_community/surveys/survey_1A_2023_04_12.csv"))
```
```{r message = FALSE}
f = list.files(path = here("data/2023/plant_community/surveys/"))%>%
  as_tibble()%>%
  mutate(data = map(value, ~read_csv(here("data/2023/plant_community/surveys/", .x), col_types = cols(quad = col_character())))) %>%
  unnest(data)
```

```{r}
survey= f %>%
  select(-value)%>%
pivot_longer(cols = c(cover1, cover2, cover3, cover4, cover5, cover6))%>%
  rename(cover = name, species = value)
```


```{r}
survey_counts <- survey %>%
  pivot_longer(cols = c(cover1count, cover2count, cover3count, cover4count, cover5count,      cover6count), values_to ="count")%>%
  mutate(name = str_replace(name, "count", ""))%>%
  filter(name == cover)%>% #is the first thing equal to the thing
   drop_na()%>%
  select(-name)
```


```{r}
species_summary = survey_counts%>%  #create new dataframe from survey_counts to make species summary with only four columns
  group_by(treatment, block, quad, species)%>% #creates new table given certain columns
  summarize(total_count = sum(count))%>% #added the counts per species to create the count total
  mutate(relative_cover = total_count/sum(total_count))%>% #relative count is the total count in eahc row divided by the sum of all the counts 
  mutate(absolute_cover = total_count/25)%>%
  rename(plot = block)
  
```

```{r}
leaf_scans = read_csv(here("data","2023","sedgwick_leaf_scans_20230421", "Summary1.csv"))
```
```{r}
leaf_dry_mass= read_csv(here("data", "2023", "traits", "leaf_dry_mass.csv"))
```
```{r}
leaf_scans_formatted = leaf_scans %>% 
  clean_names()%>%
 separate(slice, into = c("plot", "species"), sep="-") %>%
  mutate(species= str_sub(species, start= 1, end = 4), 
         species= tolower(species), 
         plot= as.numeric(plot))

```


```{r}
leaf_scan_and_dry_mass = leaf_scans_formatted %>%
  left_join(leaf_dry_mass, by= c("plot", "species"))%>%
  mutate(sla= total_area/leaf_dry_mass)
```

```{r}
species_avg_sla = leaf_scan_and_dry_mass %>%
  group_by(species) %>%
  summarize(avg_sla = mean(sla))
```


```{r}
leaf_CN = read_csv(here("data","2023","traits", "leaf_CN_20230620.csv"))
```

```{r}
leaf_CN_formatted = leaf_CN %>%
  clean_names()%>%
   rename(sample_ID = sample, mass= mass,leaf_nitrogen= nitrogen, leaf_carbon = carbon) %>%
   separate(sample_ID, into = c("plot", "species"), sep="-")%>%
  drop_na(species)%>%
  select("plot","species", "mass", "leaf_nitrogen", "leaf_carbon")

```

```{r}
leaf_CN_formatted <- leaf_CN_formatted %>%
  mutate(CNratio=leaf_carbon/leaf_nitrogen, plot= as.numeric(plot))

  #summarize(species_avg_N = sum(leaf_nitrogen)/14)
  #left_join(leaf_scan_and_dry_mass, by= c("plot", "species"))%>%
  
```

```{r}
species_avg_N <- leaf_CN_formatted%>%
  group_by(species)%>%
  summarize(species_avg_N = mean(leaf_nitrogen))
```

```{r}
species_avg_sla_and_avg_N = species_avg_N %>%
  left_join(species_avg_sla, by = c("species"))
```


```{r}
species_avg_n_and_CN <- species_avg_N %>%
  left_join(leaf_CN_formatted, by= c("species"))
```

##### This is where you need to calculate average N per species before the next chunk of code. Then you can do a left_join by species

```{r}
average_N_and_species_summary_data = species_summary %>%
   left_join(species_avg_sla_and_avg_N, by= c("species"))

```

#### You're calculating species averages in the chunk below - after you joined the leaf CN data. Use the code you wrote here but on the leaf_CN_formatted dataframe above

```{r}
#species_avg_N= leaf_CN_and_species_summary_data %>%
  #group_by(species)%>%
  #summarize(species_avg_N = mean(leaf_nitrogen))
```




```{r}
species_weighted_N_sla= average_N_and_species_summary_data %>%
  group_by(plot, treatment, quad)%>%
  summarize(average_N = weighted.mean(species_avg_N, relative_cover, na.rm = TRUE), 
            average_sla = weighted.mean(avg_sla, relative_cover, na.rm = TRUE))
 # weighted.mean = (what you're averaging first, what you're weighing it by, na.rm = drops NA before calculation)
```

```{r}
soil_moisture = read_csv(here("data/2023/soil/soil_moisture.csv")) %>% 
  mutate(soil_wet = tray_wet - tray,
         soil_dry = tray_dry - tray,
         water = soil_wet - soil_dry,
         moisture = water/soil_wet)

soil_mass = read_csv(here("data/2023/soil/soil_collected.csv")) %>% 
  mutate(soil_mass = collected_mass - rock_mass) %>% 
  left_join(soil_moisture, by = c("collection_date", "plot", "treatment", "depth")) %>% 
  mutate(soil_dry_mass = soil_mass*(1-moisture)) %>% 
  select(plot, treatment, depth, moisture, soil_dry_mass) %>% 
  mutate(depth = as.character(depth),
         plot = as.character(plot))

root_cn = read_csv(here("data/2023/roots/root_CN_20230621.csv")) %>% 
  select(sample, mass, nitrogen, carbon) %>% 
  separate(sample, into = c("plot", "depth"), sep = "_") %>% 
  filter(!is.na(depth)) %>% 
  separate(plot, into = c("plot", "treatment"), sep = 1) %>% 
  mutate(treatment = tolower(treatment),
         depth = case_when(depth == "0-20" ~ "20",
                           depth == "20-30" ~ "30"))

root_mass = read_csv(here("data/2023/roots/root_dry_matter.csv")) %>% 
  mutate(depth = as.character(depth),
         plot = as.character(plot)) %>% 
  left_join(soil_mass, by = c("plot", "treatment", "depth")) %>% 
  left_join(root_cn, by = c("plot", "treatment", "depth")) %>% 
  mutate(root_mass_cor = root_dry_mass/soil_dry_mass,
         root_c_cor = root_mass_cor * (carbon/100))
```
## You were using the wrong dataframe here: just use the species_weighted_N_sla, which is where the quadrat averages are. I think try again using the right data


Also to notice: each quadrat has two associated soil samples, because there are two depths. It's probably better to visualize the two depths separately.You can do this in your plot if you add `facet_wrap(.~depth)` to your ggplots. Google it if you want to learn more about how to use it.

```{r}
soil_sample_data <- species_weighted_N_sla %>%
  filter(quad == "ss") %>%
  mutate(treatment = tolower(treatment),
    plot= as.character(plot)) %>%
    left_join(root_mass, by = c("plot", "treatment"))
```




```{r}
ggplot(soil_sample_data, aes(x=root_mass_cor, y=average_sla)) + geom_point() + facet_wrap(.~depth)
```

```{r}
ggplot(soil_sample_data, aes(x=root_mass_cor, y=average_N)) +  geom_point() + facet_wrap(.~depth)
```

