---
title: "IRGA data for NCOS"
author: "Jacob Weverka"
date: "7/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading necessary libraries

```{r, message=FALSE}
library(tidyverse)
library(magrittr)
library(R.utils)
library(broom)
library(here)
library(janitor)
# library(BAS)
```


 <font size="4">  General setup 4</font>


Read in licor data with timestamp and date

```{r, message = FALSE}
# raw data folder
licor_raw = here::here("data/2023/SIR/licor_raw/")

files = list.files(licor_raw)
files = files[files != "desktop.ini"]

raw_data = tibble(filename = files) %>%
  mutate(contents = map(filename, ~ read_table2(file.path(licor_raw, .x), skip = 1) %>%
                          rename(time = `Time(H:M:S)`))) %>%
  mutate(timestamp_date = map(filename, ~ (str_remove(strsplit(readLines(con = file.path(licor_raw, .x), n = 1), split = " ")[[1]][1], '"'))) %>%
           lubridate::ymd(),
         timestamp_time = map(filename, ~ read_table2(file.path(licor_raw, .x), skip = 1)[1,1])) %>%
  unnest(timestamp_time)

```
Read in sequence data from lab

```{r, message=FALSE}

# excel files (CSV) with sample info folder
sequence_files = here::here("data/2023/SIR/sample_sequences/")


seqs = list.files(sequence_files)
seqs = seqs[seqs != "desktop.ini"]

date_fun = function(df){df$date[1]}

time_fun = function(df){df$time[1]}


all_seqs = tibble(filename = seqs) %>%
  mutate(sequences = map(filename, ~ read_csv(file.path(sequence_files, .x)) %>%
                          # filter(height > baseline) %>%
                          mutate(peak_number = seq.int(nrow(.))))) %>%
  mutate(timestamp_date = lubridate::mdy(map(sequences, date_fun)),
         timestamp_time = map(sequences, time_fun))

```


Add time elapsed within nested data
Filter out non-peak data
Assign peaks

```{r}
baseline = 5

run_data = raw_data %>%
  mutate(contents = map(contents,
                       ~ mutate(.x,
                                time_elapsed = seq.int(nrow(.)),
                                is_peak = `CO2(ppm)` > baseline)
                       )
         ) %>%
  unnest(contents) %>%
  filter(is_peak) %>%
  nest(-filename, -timestamp_date, -`Time(H:M:S)`) %>%
    mutate(data = 
           map(data, ~
             .x %>% mutate(peak_number = cumsum((time_elapsed - lag(time_elapsed)) > 1 & !is.na(lag(time_elapsed))) + 1
                      ) %>% 
               group_by(peak_number) %>% 
               filter(n() > 3) %>% 
               ungroup(peak_number)
           )
  )

```

Calculate area under the curve


```{r message=FALSE, warning=FALSE}
calc_area = function(df){ as.double(sum(df$rectangle, na.rm = TRUE) + sum(df$triangle, na.rm = TRUE))}

intp = run_data %>%
  mutate(data = map(data, ~ .x %>%
                      nest(-peak_number) %>%
                      mutate(data = map(data, ~ .x %>%
                                          add_row %>%
                                          mutate(
                                            rectangle = `CO2(ppm)`,
                                            triangle = abs(diff(c(baseline[1], `CO2(ppm)`[!is.na(`CO2(ppm)`)], baseline[1]))) * 0.5
                                            )
                                        )
                             ) %>%
                      mutate(area = map(data, calc_area) %>% flatten_dbl()) %>%
                      filter(area > 15) %>%
                      mutate(peak_number = seq.int(nrow(.)))
                    ),
         timestamp_date = lubridate::ymd(timestamp_date)
         )
```
Join sequence data to licor data

```{r}
# all_data =
  # inner_join(all_seqs, intp, by = "timestamp_date")
```




Joining data - be aware if you have multiple files for the same date it will produce all possible combos for that row

```{r}
all_data = as_tibble(intp) %>%
  inner_join(as_tibble(all_seqs), by = "timestamp_date") %>%
  mutate(labeled_data = map2(.x = data, .y = sequences, ~ left_join(.x, .y, by = "peak_number") %>% 
                               filter(is.na(cancel))),
         standards = map(labeled_data, ~ .x %>% drop_na(std)),
         samples = map(labeled_data, ~ .x %>% filter(is.na(std))))
```


Make standard curve

```{r message = FALSE}
add_sc = all_data %>% 
  mutate(fit = map(standards, ~ lm(std_conc ~ 0 + area, data = .)),
         results = map(fit, tidy),
         pval = map2(fit, results, ~ glance(.x) %>%
                      select(r.squared) %>%
                      bind_cols(.y$p.value)))
```


Calculate ppm values with confidence intervals for samples

```{r}
finished = add_sc %>%
  mutate(samples = map2(fit, samples,
                       ~ .y %>% bind_cols(as.data.frame(predict(.x, .y, interval = "confidence"))))
  ) %>%
  arrange(timestamp_date)
```


```{r}
weights = read_csv(here("data/2023/SIR/sir_mass.csv"),
                   col_types = cols(plot = col_character())) %>%
  mutate(treatment = toupper(treatment),
         sample_ID = paste(plot, treatment, depth, sep = "-"))
```

```{r}
data = finished %>%
  select(samples) %>%
  unnest(samples) %>%
  filter(!is.na(treatment)) %>% 
  select(-std, -std_conc, -data, -peak_number, -sample_ID) %>%
  mutate(percent_400 = replace_na(percent_400, 0),
         percent_sample = replace_na(percent_sample, 1),
         concentration = (fit - (percent_400*400))/percent_sample) %>%
  mutate(mol = (0.946353/0.082057/293.15 ) * concentration/1000000) %>%
  left_join(weights, by = c("plot", "treatment", "depth")) %>%
  group_by(sample_ID) %>%
  mutate(co2change = c(0,diff(concentration)),
         co2_cum = cumsum(co2change),
         change_adjusted = co2_cum/mass,
         time_adj = (time - min(time))/3600)
```


```{r}
# dry_mass = tibble(sample_ID = as.character(c(1,5,7,10,12)), aboveground = c(0.3412, .2948, 0.1648, .2108, .2849), root = c(.1909, .1270, .0813, .0910, .1963))
```

```{r}

ggplot(data, aes(x = time_adj, y = fit, color = sample_ID)) +
  geom_line(aes(linetype = treatment)) +
  facet_grid(. ~ depth)
```


```{r}
sum_data = data %>%
  group_by(sample_ID, plot, treatment, depth) %>%
  summarise(time_end = max(time),
            time_start = min(time),
            co2 = max(change_adjusted)) %>%
  mutate(time_elapsed = as.numeric((time_end - time_start)/3600),
         av_rate = co2/time_elapsed) %>%
  filter(!is.na(sample_ID))
```

```{r}
ggplot(sum_data, aes(x = as.factor(depth), y = av_rate, color = treatment)) +
  geom_boxplot()

# +
#   geom_smooth(aes(x = root, y = av_rate, color = NULL), method = "lm", color = "black", linetype = "dashed", se = F) +
#   theme_bw() +
#   xlab("Root Biomass (g)") +
#   ylab("Microbial Biomass SIR (mol C per hour)") +
#   scale_color_manual(name = "Treatment", values = c("darkgreen", "purple"), labels = c("Ungrazed", "Grazed")) +
#   theme(axis.text=element_text(size=12))
```

```{r}
ggplot(sum_data, aes(x = treatment, y = av_rate)) +
  geom_boxplot() +
  geom_point()





ggplot(sum_data, aes(x = treatment, y = root)) +
  geom_point()

t = lm(data = sum_data, av_rate ~ root)
summary(t)

```


```{r}
ggplot(data, aes(x = time, y = change_adjusted, color = sample_ID)) +
  geom_point() +
  geom_line()

ggplot(data, aes(x = time, y = change_adjusted, color = sample_ID)) +
  geom_col()
```



```{r}


samples = by_sample %>%
  select(sample_ID,  data, starts_data, ends_data) %>%
  #check on number of starts and ends
  mutate(nstarts = map(starts_data, nrow),
         nends = map(ends_data, nrow),
         #put starts and ends back together and arrange in order of time, assign interval number
         all = map2(starts_data, ends_data, ~ bind_rows(.x, .y) %>%
                      ungroup() %>%
                      mutate(date = lubridate::mdy(date)) %>%
                      arrange(datetime, start_end) %>%
                      mutate(interval = cumsum(start))),
         #only choose intervals with a start an an end, find change in co2 in each interval, calculate cumulative co2 emitted since start of incubation
         intervals = map(all, ~ .x %>%
                           mutate(total_time = hms::as.hms(.x$datetime - .x$datetime[1])) %>%
                           group_by(interval) %>%
                           filter(n() > 1) %>%
                           mutate(co2change = c(0, diff(mol)),
                                  timechange = (datetime - datetime[1])/86400) %>%
                           ungroup() %>%
                           mutate(total_co2 = cumsum(co2change)) %>%
                           filter(ifelse(interval == 1, start <2, start_end == "end"))
                         ))





```

```{r}

#read in soil weights, day of removal data

tubes_removal = read_csv(here("data/incubation/dpw_extraction_weights.csv"),
              col_types = cols(plot = col_factor(),
                               treatment = col_factor(),
                               depth = col_factor(),
                               sample_date = col_date(format = "%m/%d/%Y"),
                               lab_date = col_date(format = "%m/%d/%Y")))


soil_weights = read_csv(here("data/incubation/add_water.csv"),
              col_types = cols(plot = col_factor(),
                               treatment = col_factor(),
                               depth = col_factor(),
                               sample_date = col_date(format = "%m/%d/%Y"),
                               lab_date = col_date(format = "%m/%d/%Y")))


fm = read_csv(here("data/moisture_whc/csv/moisture.csv"),
              col_types = cols(date = col_date(format = "%m/%d/%Y")))
```



```{r}

fm = fm %>%
  mutate(wet_soil = tray_wet_soil - tray_weight,
         dry_soil = tray_dry_soil - tray_weight,
         moisture = wet_soil - dry_soil,
         field_moisture = moisture/dry_soil)

inc_water = soil_weights %>%
  left_join(fm, by = c("sample")) %>%
  select(sample, soil_weight, tube_ID, moisture) %>%
  mutate(water_start = moisture * soil_weight,
         soil_start = soil_weight - water_start)

```


```{r}
tubes = tubes_removal %>%
  right_join(inc_water, by = c("plot" = "sample", "tube_ID")) %>%
  rename(sample = plot) %>%
  select(sample_date, sample, tube_ID, soil_weight, soil_start)%>%
  filter(tube_ID != "X")%>%
  group_by(sample) %>%
  arrange(sample_date) %>%
  nest() %>%
  rename(tubes = data) %>%
  mutate(tubes = map(tubes, ~ .x %>%
                       mutate(total_mass = c(sum(soil_weight) - cumsum(soil_weight) + soil_weight))))
```

```{r}
air_and_soil = samples %>%
  select(sample_ID, intervals) %>%
  left_join(tubes, by = c("sample_ID" = "sample")) %>%
  mutate(as = map2(intervals, tubes, ~ left_join(.x, .y, by= c("date" = "sample_date")) %>%
                     fill(total_mass) %>%
                     fill(total_mass, .direction =  "up") %>%
                     mutate(time_elapsed = as.numeric(total_time)/86400,
                            total_time = as.numeric(total_time)/86400,
                            # time_bin = cut(total_time, breaks = c(-0.01, 0.1, 1.1, 2.2, 5.5, 7.5, 10.5, 17.5, 28.5), labels = c(0, 1, 2, 5, 7, 10, 17, 28)),
                            co2persoil = mol/total_mass,
                            co2cum = cumsum(co2persoil),
                            co2rate = co2persoil/as.numeric(timechange),
                            logco2rate = log(co2rate))
  )
  ) %>%
  select(sample_ID, as)
```

```{r}
forplot = air_and_soil %>% unnest(as)

```

```{r}
(co2cum = ggplot(forplot, aes(x = as.numeric(time_elapsed), y = co2cum, color = sample_ID)) +
  geom_line(size = 1) +
  geom_point() +
  scale_x_continuous(breaks = c(0,1,2,6,12, 19, 27, 41)) +
  ylim(0, 0.002) +
  xlab("Time (Days)") +
  ylab(expression("cumulative mol CO"[2]*" g"^"-1"*" dry soil")) )

(co2rate = ggplot(forplot %>% filter(!is.infinite(co2rate)), aes(x = as.numeric(time_elapsed), y = co2rate, color = sample_ID)) +
  geom_line(size = 1) +
  geom_point() +
   scale_x_continuous(breaks = c(0,1,2,6,12, 19, 27, 41)) +
  theme(axis.text.x = element_text(vjust = 0.5)) +
  # ggtitle("Cumulative CO2 per gram soil") +
  xlab("Time (Days)") +
  ylab(expression("mol CO"[2]*"day"^"-1"*" g"^"-1"*" dry soil")))

```